---
title: "Strikes EDA"
author: "Dave Scroggs"
date: "2023-01-27"
output: github_document
---

## Intro

This analysis and modelling process followed Dave Robinson's from [Sliced lap 1, episode 4](https://www.youtube.com/watch?v=42GyyDM3Uzg&t=4005s). This round of sliced was about predicting whether or not it would rain tomorrow in Australian cities.



```{r setup}
library(tidyverse)
library(here)
library(scales)
library(patchwork)

train <- read_csv(here("data", "MLB-pitch-data", "train.csv")) %>% 
  mutate(is_strike = factor(if_else(is_strike == 1, "strike", "no_strike")))

theme_set(theme_bw() +
            theme(plot.title = element_text(hjust = 0.5),
                  plot.subtitle = element_text(hjust = 0.5)))
```

## Skim!

The skim package is a good way to get a quick look at the columns of a table, and separates the columns by data type (categorical, numeric, logical).

```{r}
(mlb_skim <- skimr::skim(train))
```

- 17 numeric
- 3 logical
- 1 factor (calculated)
- 6 character

## Summarise function

The function below is adapted directly from DRobs notebook from [sliced ep 3](https://github.com/dgrtwo/data-screencasts/blob/c914b888df88198b67e2898216f8c6d40c072e2a/ml-practice/ep4.Rmd#L65). It's a function to summarise binomial outputs by group and gives a [credible interval](http://varianceexplained.org/r/credible_intervals_baseball/) for the percentage of strikes observed for each category.

```{r}
summarise_strikes <- function(df) {
  df %>% 
    summarise(n_strike = sum(is_strike == "strike"),
              n = n(),
              .groups = "drop") %>% 
    arrange(desc(n)) %>% 
    mutate(pct_strike = n_strike/n,
           low = qbeta(.025, n_strike + 0.5, n - n_strike + 0.5),
           high = qbeta(.975, n_strike + 0.5, n - n_strike + 0.5)) %>%
    mutate(pct = n / sum(n))
}
```

## Categorical variables

The data set has 6 categorical variables

```{r}
mlb_skim %>% 
  as_tibble() %>% 
  filter(skim_type == "character")
```

### Pitch type

```{r}

pitch_type_lookup <- tibble::tribble(
    ~pitch_description, ~pitch_type,
            "Changeup",        "CH",
           "Curveball",        "CU",
              "Cutter",        "FC",
              "Eephus",        "EP",
            "Forkball",        "FO",
  "Four-Seam Fastball",        "FA",
         "Knuckleball",        "KN",
       "Knuckle-curve",        "KC",
           "Screwball",        "SC",
              "Sinker",        "SI",
              "Slider",        "SL",
            "Splitter",        "FS",
   "Two-Seam Fastball",        "FT",
  "Sweeping curve",            "CS",
  "Four-Seam Fastball",        "FF"
  )

train %>% 
  left_join(pitch_type_lookup,
            by = "pitch_type") %>% 
  #mutate(pitch_description = fct_lump(pitch_description, prop = 0.01)) %>% 
  group_by(pitch_description) %>% 
  summarise_strikes() %>% 
  ggplot(aes(x = pct_strike, y = fct_reorder(pitch_description, pct_strike))) +
  geom_point(aes(size = n)) +
  geom_errorbar(aes(xmin = low, xmax = high), width = 0.5) +
  scale_x_continuous(labels = percent) +
  labs(x = "% strike",
       y = "Pitch type")
```

- The is a large range in the percentage of stikes thrown for each pitch type.
- Sinkers have the highest percentage of pitches that end in strikes, with a small credible interval
- Eephus(?!?), sweeping curve and knuckleballs all have large credible intervals (fewer pitches)

### Stand

```{r}
train %>% 
  group_by(stand) %>% 
  summarise_strikes() %>% 
  ggplot(aes(x = pct_strike, y = fct_reorder(stand, pct_strike))) +
  geom_point(aes(size = n)) +
  geom_errorbar(aes(xmin = low, xmax = high), width = 0.5) +
  scale_x_continuous(labels = ~percent(.x, accuracy = 1.1)) +
  labs(x = "% strike",
       y = "Batter stance")
```
- There doesn't appear to be a large difference between leftys and rightys

### Pitcher throwing arm

```{r}
train %>% 
  group_by(p_throws) %>% 
  summarise_strikes() %>% 
  ggplot(aes(x = pct_strike, y = fct_reorder(p_throws, pct_strike))) +
  geom_point(aes(size = n)) +
  geom_errorbar(aes(xmin = low, xmax = high), width = 0.5)  +
  scale_x_continuous(labels = ~scales::percent(.x, accuracy = 1.1)) +
  labs(x = "% strike",
       y = "Pitch type")
```

- As above, the doesn't appear to be a difference between lefty and righty pitchers, in isolation

### inning_topbot

```{r fig.width=10}
p1 <- train %>% 
  group_by(inning_topbot) %>% 
  summarise_strikes() %>% 
  ggplot(aes(x = pct_strike, y = fct_reorder(inning_topbot, pct_strike))) +
  geom_point() +
  geom_errorbar(aes(xmin = low, xmax = high), width = 0.5)  +
  scale_x_continuous(labels = ~scales::percent(.x, accuracy = 1.111)) +
  labs(x = "% strike",
       y = "Innings top/bottom")

p2 <- train %>% 
  filter(inning < 10) %>% 
  group_by(inning,inning_topbot) %>% 
  summarise_strikes() %>% 
  ggplot(aes(x = pct_strike, y = fct_reorder(inning_topbot, pct_strike))) +
  geom_point(aes(size = n)) +
  geom_errorbar(aes(xmin = low, xmax = high), width = 0.5) +
  scale_x_continuous(labels = ~scales::percent(.x, accuracy = 1.11)) +
  scale_size(breaks = seq(10000,20000,length.out = 5)) +
  facet_grid(inning~.) +
  labs(subtitle = "Sritke % of top and bottom of first 9 innings",
       x = "% strike",
       y = "Innings top/bottom")

p1 + p2 + patchwork::plot_layout(widths = c(1,2))
```

- Most of the credible intervals are overlapping, and percentage of strikes are similar excepts for a few innings 

### if_fielding_alignment

```{r}
train %>% 
  group_by(if_fielding_alignment) %>% 
  summarise_strikes() %>% 
  ggplot(aes(x = pct_strike, y = fct_reorder(if_fielding_alignment, pct_strike))) +
  geom_point(aes(size = n)) +
  geom_errorbar(aes(xmin = low, xmax = high), width = 0.5) +
  scale_x_continuous(labels = ~scales::percent(.x, accuracy = 1.1)) +
  labs(x = "% strike",
       y = "In-field alignment")
```
- Standard field setting has a higher % of strikes than other settings. Could this be to standard/non-standard?

### of_fielding_alignment

```{r}
train %>% 
  mutate(of_fielding_alignment = fct_lump(of_fielding_alignment, prop = 0.01)) %>% 
  group_by(of_fielding_alignment) %>% 
  summarise_strikes() %>% 
  ggplot(aes(x = pct_strike, y = fct_reorder(of_fielding_alignment, pct_strike))) +
  geom_point(aes(size = n)) +
  geom_errorbar(aes(xmin = low, xmax = high), width = 0.5) +
  labs(x = "% strike",
       y = "Out-field alignment")
```

- As above

## Logical variables

### Bases

```{r}
train %>% 
  mutate(across(on_3b:on_1b, function(col) if_else(col,"On","Off"))) %>% 
  pivot_longer(cols = on_3b:on_1b) %>% 
  group_by(name, value) %>% 
  summarise_strikes() %>% 
  ggplot(aes(x = pct_strike, y = fct_reorder(value, pct_strike))) +
  geom_point(aes(size = n)) +
  geom_errorbar(aes(xmin = low, xmax = high), width = 0.5) +
  labs(x = "% strike",
       y = "On/off base") +
  facet_grid(name ~ .)
```

- There is an apparent (small) difference in the % of strikes for all on-base situations
- For 1st base, the strike % is higher when a batter in on base, for 2nd and 3rd the opposite is true.

## Numeric variables

```{r}
mlb_skim %>% 
  as_tibble() %>% 
  filter(skim_type == "numeric") %>% 
  select(-c(character.min:logical.count))
```

### ROC

Use ROC to find out how predictive is each numeric predictor in isolation. Again, taken from DRob's [Sliced markdown](https://github.com/dgrtwo/data-screencasts/blob/c914b888df88198b67e2898216f8c6d40c072e2a/ml-practice/ep4.Rmd#L130) where he used it to find the numeric variables that were more predictive.

```{r}
train %>%
  select(is_strike, release_pos_x, release_pos_y, release_pos_z, release_speed, 
         release_spin_rate, plate_x, plate_z, spin_axis, pfx_x, pfx_z, sz_top, sz_bot) %>% 
  pivot_longer(cols = -is_strike) %>% 
  group_by(name) %>%
  yardstick::roc_auc(truth = is_strike, value, event_level = "second") %>% 
  arrange(desc(.estimate)) %>%
  mutate(name = fct_reorder(name, .estimate)) %>%
  ggplot(aes(.estimate, name)) +
  geom_point() +
  geom_vline(xintercept = .5) +
  labs(x = "AUC",
       y = "Variable",
       title = "ROC AUC of numeric variables",
       subtitle = ".5 is not predictive at all; <.5 means negatively associated with stikes, >.5 means positively associated")
```

- Pitch location at the plate (plate_x, plate_z) are the most predictive
- Spin rate, release speed, and bottom strike zone are also predictive

## Release

Release position is the position of the ball when it is pitched, relative to a top-down view of the pitcher.

```{r}
p1 <- train %>% 
  ggplot(aes(x = release_pos_x, y = release_pos_z, col = p_throws)) +
  geom_density_2d() +
  scale_colour_manual(values = c("L" = "#1874CD", "R" = "#FF3030")) +
  labs(x = "Release position x", 
       y = "Release position y") +
  facet_grid(is_strike~.)

p2 <- train %>% 
  mutate(release_pos_x = if_else(p_throws == "L", -release_pos_x, release_pos_x)) %>% 
  ggplot(aes(x = release_pos_x, y = release_pos_z, col = p_throws)) +
  geom_density_2d() +
  scale_colour_manual(values = c("L" = "#1874CD", "R" = "#FF3030")) +
  labs(x = "Release position x", 
       y = "Release position y") +
  facet_grid(is_strike~.)

p1 + p2 + 
  plot_layout(guides = "collect") +
  plot_annotation(
    title = "Pitcher relative position, right plot is adjusted for left/right pitcher",
    subtitle = "The pitch x release is flipped if pitcher is left handed",
    theme = theme(
      plot.title = element_text(hjust = 0.5),
      plot.subtitle = element_text(hjust = 0.5)
    )
  )

```

- Not sure if this is a good idea, as it would removed the left/right relationship between the batter and pitcher, but maybe everything is relative - ie. it matters that the ball starts on the batters outer side, rather than left/right dynamics. 

## Engineered categoricals

This section contains numeric variables that have been converted to categorical types to look at the effect on strike percentage.


#### Combining on-bases

```{r}
train %>%
  select(uid, on_1b, on_2b, on_3b, is_strike) %>% 
  pivot_longer(cols = c(on_1b, on_2b, on_3b)) %>% 
  mutate(base = str_extract(name, "\\d")) %>% 
  filter(value) %>% 
# This probably should have been done with a long case_when but I cbf
{bind_rows(., filter(
  (train %>% transmute(uid, is_strike, name = "", value = TRUE, base = "0")),
  !uid %in% unique(.$uid)))} %>% 
  group_by(uid, is_strike) %>% 
  summarise(bases = paste(base, collapse = ", "), .groups = "drop") %>%
  group_by(bases) %>% 
  summarise_strikes() %>% 
  ggplot(aes(x = pct_strike, y = fct_reorder(bases, pct_strike))) +
  geom_point(aes(size = n)) +
  geom_errorbar(aes(xmin = low, xmax = high), linewidth = 0.3, width = 0.2) +
  labs(title = "Batters on base when at bat",
       x = "% strike",
       y = "On/off base")
```

- Zero batters on-base is the most common situation
- Strike % is higher when there is only a runner on 1st
- There is a smaller strike % for the less frequent on-base situations, but they have large credible intervals. 


### Count

The count in baseball is the current number of balls and strikes on the batter.

### Balls/strikes

```{r}
train %>% 
  filter(balls < 4) %>% 
  group_by(strikes, balls) %>% 
  summarise_strikes() %>% 
  ggplot(aes(x = pct_strike, y = fct_reorder(factor(balls), pct_strike))) +
  geom_point(aes(size = n)) +
  geom_errorbar(aes(xmin = low, xmax = high), width = 0.5) +
  facet_grid(paste(strikes,"strike(s)") ~ .) +
  labs(title = "Strike percentage for ball/strike status",
       x = "% strikes", 
       y = "balls")
```

```{r}
train %>% 
  filter(balls < 4) %>% 
  group_by(balls, strikes) %>% 
  summarise_strikes() %>% 
  ggplot(aes(x = balls, y = strikes, fill = pct_strike)) +
  geom_tile() +
  geom_text(aes(label = scales::percent(pct_strike, accuracy = 1.1)), col = "white")
```

- Strike % is lower when the ball count is lower, or when the strike count is higher

### In zone

```{r}
# Horizontal strike-zone from mid-point of home plate
horiz_sz = (17/2 + 2.9)/12

train %>% 
  # Find if pitch was in/out of strike zone
  mutate(inZone = if_else(between(plate_x,-horiz_sz, horiz_sz) & between(plate_z,sz_bot, sz_top), "inZone","outZone")) %>% 
  group_by(inZone) %>% 
  summarise_strikes() %>% 
  ggplot(aes(x = pct_strike, y = fct_reorder(inZone, pct_strike))) +
  geom_point() +
  geom_errorbar(aes(xmin = low, xmax = high), width = 0.5) +
  labs(x = "% strikes", 
       y = "In/Out strike zone")
```

- Unsurprisingly the difference between the percentage of strikes in and out of the stike zone is large.

### Outs when up

```{r}
train %>%
  group_by(outs_when_up) %>% 
  summarise_strikes() %>% 
  ggplot(aes(x = pct_strike, y = fct_reorder(factor(outs_when_up), pct_strike))) +
  geom_point(size = 3) +
  geom_errorbar(aes(xmin = low, xmax = high), width = 0.5) +
  labs(x = "% strikes", 
       y = "Number of outs")
```
- Strike % is higher when there are 0 outs
- Removed point size for number of counts because small differences in the counts we being shown as large size differences in the plot

## Multi-variate

### Strike zone

Horizontal strike zone, around the mid-point of home plate = 17 inches/2 + 2.9 inches (diameter of baseball)

```{r}
left_sz = (17/2 + 2.9)/12

train %>% 
  ggplot(aes(x = plate_x, y = plate_z)) +
  geom_bin_2d(binwidth = c(0.1,0.1)) +
  geom_vline(xintercept = left_sz) +
  geom_vline(xintercept = -left_sz) +
  facet_wrap(~is_strike) +
  labs(caption = "Horizontal strike zone shown as vertical lines")
```

## Pitch location by pitch type

```{r}
train %>% 
  left_join(pitch_type_lookup,
            by = join_by(pitch_type)) %>% 
  ggplot(aes(x = pfx_x, y = pfx_z)) +
  geom_bin2d() +
  facet_grid(p_throws~pitch_description)
```

## Numeric variable feature engineering

```{r}
train %>% 
  mutate(plate_x_eng = if_else(stand == "L", -plate_x, plate_x),
         release_pos_x_eng = if_else(p_throws == "L", -release_pos_x, release_pos_x),
         inZone_low_eng = plate_z - sz_bot,
         inZone_high_eng = plate_z - sz_top) %>% 
  select(is_strike, release_pos_x, release_pos_y, release_pos_z, release_speed, 
         release_spin_rate, plate_x, plate_z, spin_axis, pfx_x, pfx_z, sz_top, sz_bot,
         plate_x_eng, release_pos_x_eng, inZone_low_eng, inZone_high_eng) %>% 
  pivot_longer(cols = -is_strike) %>% 
  group_by(name) %>%
  yardstick::roc_auc(truth = is_strike, value, event_level = "second") %>% 
  arrange(desc(.estimate)) %>% 
  mutate(name = fct_reorder(name, .estimate),
         engineered = if_else(str_detect(name, "_eng"),"engineered", "standard")) %>%
  ggplot(aes(.estimate, name)) +
  geom_point(aes(col = engineered)) +
  scale_colour_manual(values = c("engineered" = "red", "standard" = "black")) +
  geom_vline(xintercept = .5) +
  labs(x = "AUC",
       y = "Variable",
       title = "ROC AUC of numeric variables",
       colour = "Variables",
       subtitle = ".5 is not predictive at all; <.5 means negatively associated with stikes, >.5 means positively associated")
```
 - Some of the engineered features show improved differentiation in the numeric variables, especially when considering left/right batter position in the x plane.


## Question

- When is it advisable/not advisable to engineer features to infer a play situation - ie. the count or base situation, rather than have each feature available for split. Is it preferable to let the algorithm decide whether to split say for left handed batters, plate_x > X, right handed batters plate_x < Y etc?