---
title: "First attempt at NWVS"
output: github_document
---


I couldn't play along at the time because it was the middle of the day here, plut this took me longer than 2 hours.

## NWVS - predict the build year of houses based upon their features

```{r}
library(tidyverse)
library(tidymodels)
library(here)

train <- read_csv(here("data/NWVS/train.csv"))
test <- read_csv(here("data/NWVS/test.csv"))

houses_cv <- train %>% 
  vfold_cv(v = 10)
```

## SKIM!

```{r}
(skimobj <- skimr::skim(train))
```


## EDA

### Predictor

```{r}
train %>% 
  count(yearBuilt) %>% 
  ggplot(aes(x = yearBuilt, y = n)) +
  geom_col(fill = "dodgerblue", col = "black") + 
  geom_curve(x = 1875, y = 400, xend = min(train$yearBuilt), yend = 20,
             arrow = arrow(length = unit(0.03, "npc"), type = "closed")) +
  annotate("text", x = 1875, y = 400, label = "Earliest house - 1860", hjust = 0) +
  geom_curve(x = 2000, y = 600, xend = max(train$yearBuilt), yend = 20, curvature = -0.5,
             arrow = arrow(length = unit(0.03, "npc"), type = "closed")) +
  annotate("text", x = 2000, y = 600, label = "Latest house - 2021", hjust = 1) +
  labs(x = "Year built",
       y = "# homes") +
  theme_bw()
```
### Categorical predictors

```{r}
categ_cols <- function(df, col){
  df %>% 
    count({{col}}) %>% 
    mutate("{{col}}" := fct_reorder({{col}}, n),
           pct = scales::percent(n/sum(n), accuracy = 1.1)) %>% 
    ggplot(aes(x = n, y = {{col}})) +
    geom_col(fill = "dodgerblue", colour = "black") +
    geom_text(aes(label = pct))
}
```


```{r}
skimobj %>% 
  filter(skim_type == "character") %>% 
  as_tibble()
```

#### Home type

```{r}
train %>% 
  categ_cols(homeType)
```

- Few categories, most of the rows are of one type - Single_family

#### architecturalStyle

```{r}
train %>% 
  categ_cols(architecturalStyle)
```

- Largest group is missing data (lots of work to impute values today!)
- Most of the rows of in a few different categories (lump smaller groups)

#### constructionMaterials

```{r}
train %>% 
  categ_cols(constructionMaterials)
```
- Significant missing data (lots of work to impute values today!)
- Again, most of the rows of in a few different categories (lump smaller groups)

Materials are probably a good indicator for year built

```{r}
train %>% 
  mutate(constructionMaterials = fct_lump(as.character(constructionMaterials), n = 10, other_level = "other")) %>%
  count(yearBuilt, constructionMaterials) %>% 
  ggplot(aes(x = yearBuilt, y = n)) +
  geom_col(col = "black", fill = "dodgerblue") +
  facet_wrap(~constructionMaterials, scales = "free_y", ncol = 2)
```

 - Looks like certain materials were used in different eras, but gotta be careful not to overfit
 - 10 levels seems a good start

### Numerical predictors

```{r}
skimobj %>% 
  filter(skim_type == "numeric") %>% 
  as_tibble() %>% 
  select(skim_type:complete_rate) %>% 
  knitr::kable()
```

 - Some of these could be interpreted as ordinal/categorical maybe? Bedrooms, bathrooms, stories?
 
#### lastSoldPrice

```{r}
train %>% 
  group_by(yearBuilt) %>% 
  summarise(median = median(lastSoldPrice, na.rm = TRUE)) %>% 
  ggplot(aes(x = yearBuilt, y = median)) +
  geom_point() +
  labs(y = "Median last sold price")
```

#### Lotsize

```{r}
train %>% 
  group_by(yearBuilt) %>% 
  summarise(median = median(lotSize, na.rm = TRUE)) %>% 
  ggplot(aes(x = yearBuilt, y = median)) +
  geom_point() +
  labs(y = "Median lot size")
```


#### Potential categoricals

##### Bedrooms

```{r}
train %>% 
  mutate(bedrooms = fct_lump(as.character(bedrooms), n = 6)) %>%
  count(yearBuilt, bedrooms) %>% 
  ggplot(aes(x = yearBuilt, y = n)) +
  geom_col(col = "black", fill = "dodgerblue") +
  facet_grid(bedrooms~., scales = "free_y")
```

##### Bathrooms

```{r}
train %>% 
  mutate(bathrooms = as.character(bathrooms)) %>% 
  categ_cols(bathrooms)
```
 - Lump this and convert to categorical


##### Stories

```{r}
train %>% 
  mutate(stories = as.character(stories)) %>% 
  categ_cols(stories)
```

##### hasAttachedGarage

```{r}
train %>% 
  mutate(hasAttachedGarage = as.character(hasAttachedGarage)) %>% 
  categ_cols(hasAttachedGarage)
```

 - Essentially a dummy var

##### hasFireplace

```{r}
train %>% 
  mutate(hasFireplace = as.character(hasFireplace)) %>% 
  categ_cols(hasFireplace)
```

```{r}
train %>% 
  ggplot(aes(x = yearBuilt, fill = factor(hasFireplace))) +
  geom_bar(position = position_fill())
  
```


##### hasView

```{r}
train %>% 
  mutate(hasView = as.character(hasView)) %>% 
  categ_cols(hasView)
```

### Missing values

```{r}
skimobj %>% 
  filter(n_missing > 0) %>% 
  as_tibble() %>% 
  select(skim_type:complete_rate) %>% 
  arrange(-n_missing)
```


## Model build


### Recipes

 - Impute missing
    - Categorical: architecturalStyle, constructionMaterials
 - Lump smaller groups: architecturalStyle, constructionMaterials

From the recipes [website](https://recipes.tidymodels.org/articles/Ordering.html)
Order
1. Impute
2. Handle factor levels
3. Individual transformations for skewness and other issues
4. Discretize (if needed and if you have no other choice)
5. Create dummy variables
6. Create interactions
7. Normalization steps (center, scale, range, etc)
8. Multivariate transformation (e.g. PCA, spatial sign, etc)

```{r}
tree_recipe <- 
  recipe(formula = yearBuilt ~ ., data = train) %>% 
  step_zv(all_predictors()) %>% 
  update_role(houseID, new_role = "id") %>% 
  # I don't know the best way to do this!
  step_impute_median(stories) %>% 
  # impute missing data - hasAttachedGarage, hasFireplace
  step_impute_median(hasFireplace, hasAttachedGarage) %>% 
  # architecturalStyle, constructionMaterials,
  step_impute_mode(architecturalStyle, constructionMaterials) %>% 
  # lotSize,
  step_impute_knn(lotSize, impute_with = imp_vars(lastSoldPrice, latitude, longitude)) %>%
  step_mutate(bathrooms = round(bathrooms)) %>% 
  step_num2factor(stories, levels = as.character(1:5)) %>% 
  step_other(architecturalStyle, constructionMaterials, stories, threshold = 0.05) %>% 
  step_unknown(all_nominal_predictors()) %>%
  step_dummy(all_nominal_predictors())

  
 tree_recipe %>% prep %>% bake(train)

```


### Model spec

 - Fit a model with some default values
 
```{r}
ranger_spec <- 
  rand_forest(mtry = 10, min_n = 15, trees = 1000) %>% 
  set_mode("regression") %>% 
  set_engine("ranger")

xgboost_spec <- 
  boost_tree(trees = 1250, min_n = 15, tree_depth = 10, learn_rate = 0.02) %>% 
  set_mode("regression") %>% 
  set_engine("xgboost")
```

### Fit resamples

```{r}
# metrics
mset <- metric_set(rmse)

control <- control_grid(verbose = TRUE, save_pred = TRUE)

model_trials <- workflow_set(preproc = list(tree_rec = tree_recipe),
             models = list(rf = ranger_spec, xgb = xgboost_spec),
             cross = TRUE)

model_resamples <- workflow_map(
  model_trials,
  "fit_resamples",
  resamples = houses_cv,
  metrics = mset,
  control = control,
  verbose = TRUE
)
```

```{r}
model_resamples %>% 
  collect_metrics()
```


#### Check residuals

```{r}
model_resamples %>% 
  collect_predictions() %>% 
  ggplot(aes(x = yearBuilt, y = .pred)) +
  geom_point(alpha = 0.05) +
  geom_abline(col = "blue")
```
 - The earlier and later years aren't getting predicted well

### Model tuning

- TBA

#### Model workflows

##### Random Forest

```{r}
ranger_spec <- 
  rand_forest(mtry = tune(), min_n = tune(), trees = 1000) %>% 
  set_mode("regression") %>% 
  set_engine("ranger")

xgboost_spec <- 
  boost_tree(trees = tune(), min_n = tune(), tree_depth = tune(), learn_rate = tune()) %>% 
  set_mode("regression") %>% 
  set_engine("xgboost")
```


```{r}
set.seed(65652)
ranger_tune <-
  tune_grid(ranger_workflow, resamples = stop("add your rsample object"), grid = stop("add number of candidate points"))

xgboost_tune <-
  tune_grid(xgboost_workflow, resamples = stop("add your rsample object"), grid = stop("add number of candidate points"))
```

##### XGBoost


###### Step 1: Fix learning rate and number of estimators for tuning tree-based parameters
```{r}
xgboost_spec1 <- 
  boost_tree(trees = tune(), min_n = 10, tree_depth = 5, learn_rate = 0.1, loss_reduction = 0) %>% 
  set_mode("regression") %>% 
  set_engine("xgboost")

grid_step1 <- grid_regular(trees(range = c(100,1000)), levels = 10)

xgboost_tune1 <-
  tune_grid(
    object = xgboost_spec1,
    preprocessor = tree_recipe,
    resamples = houses_cv,
    grid = grid_step1,
    metrics = mset,
    control = control
  )

xgboost_tune %>% 
  autoplot()
```

##### Step 2

- Trees: 700

2. Tune tree-specific parameters ( max_depth, min_child_weight, gamma, subsample, colsample_bytree) for decided learning rate and number of trees. Note that we can choose different parameters to define a tree and I’ll take up an example here.

```{r}
xgboost_spec2 <- 
  boost_tree(learn_rate = 0.02,
    trees = 1000,
    mtry = tune(),
    min_n = 10,    
    tree_depth = tune()) %>% 
  set_mode("regression") %>% 
  set_engine("xgboost")

grid_step2 <- grid_latin_hypercube(
    mtry(range = c(10, 17)),
    tree_depth(range = c(8, 20)),size = 30)

xgboost_tune2 <-
  tune_grid(
    object = xgboost_spec2,
    preprocessor = tree_recipe,
    resamples = houses_cv,
    grid = grid_step2,
    metrics = mset,
    control = control
  )

xgboost_tune2 %>% 
  autoplot()

xgboost_tune2 %>% select_best()
```

### DGRTWO Method

- Pick out the numeric variables, or the low-cardinality categorical ones
- Mean-impute missing values, if any


```{r}
# Recipe

dgr_recipe <- 
  recipe(formula = yearBuilt ~ ., data = train) %>% 
  step_zv(all_predictors()) %>% 
  update_role(houseID, new_role = "id") %>% 
  step_impute_mean(stories, hasFireplace, hasAttachedGarage, lotSize) %>% 
  step_impute_mode(architecturalStyle, constructionMaterials) %>% 
  step_mutate(bathrooms = round(bathrooms)) %>% 
  step_num2factor(stories, levels = as.character(1:5)) %>% 
  step_other(architecturalStyle, constructionMaterials, stories, threshold = 0.05) %>% 
  step_unknown(all_nominal_predictors()) %>%
  step_dummy(all_nominal_predictors())
```

- Fit an xgboost boosted trees model, with learn_rate = .02, mtry varying from 2-4, and trees varying from 300 to 800
- Plot a learning curve (showing the performance as the number of trees varies), and use it to adjust mtry and trees if needed.

```{r}
xgboost_spec_dgr <- 
  boost_tree(learn_rate = 0.02,
    trees = tune(),
    mtry = tune(),
    tree_depth = tune()) %>% 
  set_mode("regression") %>% 
  set_engine("xgboost")

grid_step_dgr <- grid_latin_hypercube(
  trees(range = c(300,800)),
    mtry(range = c(2, 10)),
    tree_depth(range = c(5, 20)),size = 30)

xgboost_tune_dgr <-
  tune_grid(
    object = xgboost_spec_dgr,
    preprocessor = dgr_recipe,
    resamples = houses_cv,
    grid = grid_step_dgr,
    metrics = mset,
    control = control
  )

xgboost_tune_dgr %>% collect_metrics() %>% filter(trees > 700) %>% ggplot(aes(tree_depth, mtry, col = trees)) +geom_point()
  autoplot()

xgboost_tune_dgr %>% 
  select_best()
```


#### Check fit

```{r}
xgb_resamples <- workflow(preprocessor = tree_recipe,
                           spec = finalize_model(xgboost_spec2, select_best(xgboost_tune2))) %>% 
  fit_resamples(resamples = houses_cv,
  metrics = mset,
  control = control)

xgb_resamples %>% 
  collect_metrics()
```



### Final fit

```{r}

# Early model
prod_model <- workflow() %>%
  add_recipe(tree_recipe) %>%
  add_model(ranger_spec) %>%
  fit(train)

txgb_fit_final <- workflow(preprocessor = tree_recipe,
                           spec = boost_tree(learn_rate = 0.02,
    trees = 1000,
    mtry = 15,
    min_n = 10,    
    tree_depth = 18) %>% 
  set_mode("regression") %>% 
  set_engine("xgboost")) %>% 
  fit(train)
```



### Create submission

```{r}
txgb_fit_final %>% 
    augment(test) %>%
    select(houseID, yearBuilt = .pred) %>% 
  write_csv("../data/NWVS/fine_tuned_xgb_submission.csv")
```

