---
title: "Sliced Ep 12"
subtitle: "U.S. Small Business Administration (SBA) loans"
output: html_document
date: '2022-06-22'
---

```{r}
library(tidyverse)
library(skimr)
library(scales)
library(here)

SBA_train <- read_csv(here("data/sliced-s01e12/train.csv"))

SBA_train
```

## Quick overview

```{r}
skim(SBA_train)
```
- Very little missing data
- 6 character variables
- 14 numeric
- prediction variable = default_amont


## Default amount

```{r}
SBA_train %>% 
  filter(default_amount > 0) %>% 
  ggplot(aes(default_amount)) +
  geom_histogram() +
  scale_x_log10()
```
- Distribution of default payments (if made) is log-normal

```{r}
SBA_train %>% 
  count(default_amount > 0) %>% 
  mutate(pct = (n/sum(n)) %>% percent())
```

3/4 of the rows are do not have a default amount paid

## Character columns

```{r}
SBA_train %>% 
  select_if(is.character) %>% 
  colnames()
```

```{r}
SBA_train %>% 
  select_if(is.character) %>% 
  summarise_all(~unique(.x) %>% length())

```

### State


```{r}
SBA_train %>% 
  mutate(tf = default_amount > 0) %>% 
  group_by(State,tf) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n)) %>% 
  ggplot(aes(x = pct,y = fct_reorder(State,pct,min),fill = tf)) + 
  geom_col(position = "stack") +
  geom_text(aes(label = percent(pct,1)),position = position_stack(vjust = 0.5)) +
  labs(x = "Percent default",
       y = "State") +
  theme(axis.text.x = element_text(angle = 45,hjust = 1),
        legend.position = "bottom")
```

```{r}
SBA_train %>% 
  filter(default_amount > 0) %>% 
  mutate(log_default_amt = log(default_amount)) %>% 
  group_by(State) %>% 
  summarise(default_mean = mean(log_default_amt),
            default_median = median(log_default_amt)) %>% 
  ggplot(aes(y = fct_reorder(State,default_mean))) + 
  geom_point(aes(x = default_mean), col = "red") +
  geom_point(aes(x = default_median), col = "blue") +
  labs(x = "log mean/median default",
       y = "State") +
  theme(axis.text.x = element_text(angle = 45,hjust = 1),
        legend.position = "bottom")
```

```{r}
library(usmap)
SBA_train %>% 
  rename(state = State) %>% 
  count(state, wt = default_amount,sort = T) %>% 
plot_usmap(data = .,regions = "state",values = "n", color = "white") + 
  scale_fill_continuous(name = "Default amount", label = scales::comma) + 
  theme(legend.position = "right")

```
### Sector

```{r}
SBA_train %>% 
  count(Sector,wt = default_amount,sort = T) %>% 
  ggplot(aes(x = n, y = fct_reorder(Sector,n))) + 
  geom_col() +
  labs(x = "Total default amount ($)",
       y = "Sector") +
  scale_x_continuous(labels = dollar) +
  theme(axis.text.x = element_text(angle = 45,hjust = 1))
```


```{r}
SBA_train %>% 
  mutate(tf = default_amount > 0) %>% 
  group_by(Sector,tf) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n)) %>% 
  ggplot(aes(x = pct,y = fct_reorder(Sector,pct,min),fill = tf)) + 
  geom_col(position = "stack") +
  geom_text(aes(label = percent(pct,1)),position = position_stack(vjust = 0.5)) +
  labs(x = "Percent default",
       y = "Sector") +
  theme(axis.text.x = element_text(angle = 45,hjust = 1),
        legend.position = "bottom")
```

```{r}
SBA_train %>% 
  filter(default_amount > 0) %>% 
  mutate(log_default_amt = log(default_amount)) %>% 
  group_by(Sector) %>% 
  summarise(default_mean = mean(log_default_amt),
            default_median = median(log_default_amt)) %>% 
  ggplot(aes(y = fct_reorder(Sector,default_mean))) + 
  geom_point(aes(x = default_mean), col = "red") +
  geom_point(aes(x = default_median), col = "blue") +
  labs(x = "Percent default",
       y = "Sector") +
  theme(axis.text.x = element_text(angle = 45,hjust = 1),
        legend.position = "bottom")
```

```{r}
SBA_train %>% 
  #filter(default_amount > 0) %>% 
  filter(default_amount > 0,str_detect(Sector, "Public|Real")) %>% 
  mutate(log_default_amt = log(default_amount)) %>%
  ggplot(aes(log_default_amt,col = Sector)) +
  geom_density() +
  theme(legend.position = "bottom")
```

Looks like the "Public administration sector" is bi-modal. Could be worth separating into high/low amounts.

### Banks

```{r}
SBA_train %>% 
  filter(default_amount > 0) %>% 
  mutate(log_default_amt = log(default_amount)) %>% 
  group_by(Bank) %>% 
  summarise(default_mean = mean(log_default_amt),
            default_median = median(log_default_amt)) %>% 
  ggplot(aes(y = fct_reorder(Bank,default_mean))) + 
  geom_point(aes(x = default_mean), col = "red") +
  geom_point(aes(x = default_median), col = "blue") +
  labs(x = "Percent default",
       y = "Bank") +
  theme(axis.text.x = element_text(angle = 45,hjust = 1),
        legend.position = "bottom")
```
```{r}
SBA_train %>% 
  mutate(log_default_amt = log(default_amount),
         Bank = fct_lump_prop(Bank,0.003)) %>%
  group_by(Bank,tf = default_amount > 0) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n)) %>% 
  filter(tf) %>% 
  ggplot(aes(x = pct, y = fct_reorder(Bank,pct))) + 
  geom_point(aes(size = n)) +
  labs(x = "Percent default",
       y = "Bank") +
  theme(axis.text.x = element_text(angle = 45,hjust = 1),
        legend.position = "bottom")

SBA_train %>% 
  filter(default_amount > 0) %>% 
  mutate(log_default_amt = log(default_amount)) %>% 
  group_by(Bank,log_default_amt) %>% 
  summarise(n = n()) %>% 
  mutate(pct = n/sum(n)) %>% 
  arrange(-pct)
```

## Numerical columns

```{r}
SBA_train %>% 
  select_if(is.numeric)
```

```{r}
SBA_train %>% 
  filter(default_amount > 0) %>% 
  group_by(ApprovalFY) %>% 
  summarise(default_mean = mean(default_amount)) %>% 
  ggplot(aes(ApprovalFY,default_mean)) +
  geom_col() +
  scale_y_continuous(labels = dollar)
```

```{r}
SBA_train %>% 
  filter(default_amount > 0) %>% 
  mutate(Sector = fct_lump(Sector,10)) %>% 
  count(Sector,ApprovalFY, wt = default_amount) %>% 
  ggplot(aes(ApprovalFY,n)) +
  geom_col() +
  scale_y_continuous(labels = dollar) +
  facet_wrap(~Sector)
```

```{r}
SBA_train %>% 
  ggplot(aes(DisbursementGross,SBA_Appv)) +
  geom_point(alpha = 0.3) +
  geom_abline(slope = 1) +
  geom_abline(slope = 0.9) +
  geom_abline(slope = 0.1) +
  geom_abline(slope = 0.75) +
  geom_hline(yintercept = 1E6) +
  geom_hline(yintercept = 75E4)
```

```{r}
SBA_train %>% 
  ggplot(aes(DisbursementGross,default_amount)) +
  geom_point() +
  geom_abline()
```

Definietly correlated

```{r}
SBA_train %>% 
  filter(Name == "SUBWAY")
```

```{r}
SBA_train %>% 
  ggplot(aes(NewExist,fill = default_amount > 0)) +
  geom_bar(position = "fill")
```

```{r}
SBA_train %>%
  filter(NewExist %in% 1:2) %>%
  ggplot(aes(x = log(default_amount),fill = factor(NewExist))) +
  geom_histogram(binwidth = 0.1) +
  facet_wrap(~NewExist,scales = "free_y")
```

